Location and Scale-Invariant Power Transformations for Transforming Data to Normality

### 2024-10-18 manuscript_latex/manuscript.tex
Make figures fit in width.
figures were created for pages with smaller margins than elsarticle
Set width of figures.
Added abstract and highlights.

\begin{abstract}

\end{abstract}

@@

\begin{highlights}
\item Research highlight 1
\item Research highlight 2
\end{highlights}

@@

%% keywords here, in the form: keyword \sep keyword

@@

{\centering \includegraphics{figure_2}

@@

{\centering \includegraphics{figure_3}

@@

{\centering \includegraphics{figure_4}

@@

{\centering \includegraphics{figure_5}

@@

{\centering \includegraphics{figure_6}

@@

{\centering \includegraphics{figure_7}

@@

{\centering \includegraphics{figure_8}

@@

{\centering \includegraphics{figure_9}

@@

{\centering \includegraphics{figure_10}

@@

{\centering \includegraphics{figure_appendix_1}

@@

{\centering \includegraphics{figure_appendix_2}

@@

{\centering \includegraphics{figure_appendix_3}

@@

{\centering \includegraphics{figure_appendix_4}

@@

\midrule
Box-Cox                  & 0.047 & 0.073 & 0.090 & 0.106 & 0.126 & 0.140 & 0.188 \\
Box-Cox (no outlier)     & 0.043 & 0.065 & 0.079 & 0.092 & 0.106 & 0.116 & 0.155 \\
Yeo-Johnson              & 0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
Yeo-Johnson (no outlier) & 0.041 & 0.061 & 0.074 & 0.085 & 0.099 & 0.109 & 0.139 \\
normal distr.            & 0.039 & 0.066 & 0.083 & 0.097 & 0.117 & 0.132 & 0.174 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

@@

\toprule
feature & none & conventional & conv. (z-score norm.) & conv. (rob. scaling) & invariant \\

>>>>

\begin{abstract}
Power transformations are used to stabilize variance and achieve
normality of features, especially in methods assuming normal 
distributions such as ANOVA and linear discriminant analysis. 
However, the commonly used Box-Cox and Yeo-Johnson power transformation 
methods are sensitive to the location, scale, and presence of 
outliers in the data.

Here we present location- and scale-invariant Box-Cox and Yeo-Johnson
transformations to mitigate these issues. We derive maximum likelihood
estimation criteria for optimizing transformation parameters and propose
robust adaptations that reduce the influence of outliers. 
We also introduce an empirical test for assessing central normality
of transformed features.

In simulations and real-world datasets, robust location- and scale-invariant 
transformations outperform conventional variants, resulting in better 
transformations to central normality. In a machine learning experiment
with 232 datasets with numerical features, integrating robust 
ocation- and scale-invariant power transformations into an automated data 
processing and machine learning pipeline did not result in a meaningful
improvement or detriment in model performance compared to conventional variants.

In conclusion, robust location- and scale-invariant power
transformations can replace conventional variants.
\end{abstract}

@@

\begin{highlights}
\item Location and scale of feature distributions may cause wrong power transformations.
\item We derived location- and scale-invariant Box-Cox and Yeo-Johnson transformations.
\item Robust power transformations succesfully reduced negative effects of outliers.
\item New invariant robust power transformations can replace conventional variants.
\end{highlights}

@@

power transformation \sep invariant transformation \sep robust transformation \sep normality test
%% keywords here, in the form: keyword \sep keyword

@@

{\centering \includegraphics[width=1\linewidth]{figure_2}

@@

{\centering \includegraphics[width=1\linewidth]{figure_3}

@@

{\centering \includegraphics[width=1\linewidth]{figure_4}

@@

{\centering \includegraphics[width=1\linewidth]{figure_5}

@@

{\centering \includegraphics[width=1\linewidth]{figure_6}

@@

{\centering \includegraphics[width=1\linewidth]{figure_7}

@@

{\centering \includegraphics[width=1\linewidth]{figure_8}

@@

{\centering \includegraphics[width=1\linewidth]{figure_9}

@@

{\centering \includegraphics[width=1\linewidth]{figure_10}

@@

{\centering \includegraphics[width=1\linewidth]{figure_appendix_1}

@@

{\centering \includegraphics[width=1\linewidth]{figure_appendix_2}

@@

{\centering \includegraphics[width=1\linewidth]{figure_appendix_3}

@@

{\centering \includegraphics[width=1\linewidth]{figure_appendix_4}

@@

\midrule
Box-Cox                  & 0.047 & 0.073 & 0.090 & 0.106 & 0.126 & 0.140 & 0.188 \\
Box-Cox (no outlier)     & 0.043 & 0.065 & 0.079 & 0.092 & 0.106 & 0.116 & 0.155 \\
Yeo-Johnson              & 0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
Yeo-Johnson (no outlier) & 0.041 & 0.061 & 0.074 & 0.085 & 0.099 & 0.109 & 0.139 \\
normal distr.            & 0.039 & 0.066 & 0.083 & 0.097 & 0.117 & 0.132 & 0.174 \\
\bottomrule

\end{tabular}
\end{center}
\end{table}

@@

\toprule
feature & none & conv. & conv. (z-score norm.) & conv. (rob. scaling) & invariant \\
